---
title: "african crises"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
require(tidyverse)
require(parsnip)
require(tidymodels)

dat <- rio::import(here::here("data", "african_crises.csv"))
```

caseA number which denotes a specific country
cc3A three letter country code
countryThe name of the country
yearThe year of the observation
systemic_crisis"0" means that no systemic crisis occurred in the year and "1" means that a systemic crisis occurred in the year.
exch_usdThe exchange rate of the country vis-a-vis the USD
domestic_debt_in_default"0" means that no sovereign domestic debt default occurred in the year and "1" means that a sovereign domestic debt default occurred in the year
sovereign_external_debt_default"0" means that no sovereign external debt default occurred in the year and "1" means that a sovereign external debt default occurred in the year
gdp_weighted_defaultThe total debt in default vis-a-vis the GDP
inflation_annual_cpiThe annual CPI Inflation rate
independence"0" means "no independence" and "1" means "independence"
currency_crises"0" means that no currency crisis occurred in the year and "1" means that a currency crisis occurred in the year
inflation_crises"0" means that no inflation crisis occurred in the year and "1" means that an inflation crisis occurred in the year
banking_crisis"no_crisis" means that no banking crisis occurred in the year and "crisis" means that a banking crisis occurred in the year

```{r}
dat %>% count(country)

dat %>% count(year)

dat %>% 
  ggplot(aes(year)) +
  geom_histogram(binwidth = 1)

cpi_by_country <- dat %>% 
  group_by(country) %>% 
  summarize_at(vars(exch_usd, 
                    domestic_debt_in_default,
                    sovereign_external_debt_default,
                    gdp_weighted_default),
               mean)

cpi_by_country %>% 
  mutate(country = fct_reorder(country, desc(sovereign_external_debt_default))) %>% 
  ggplot(aes(country, sovereign_external_debt_default)) +
  geom_bar(stat = "identity")


dat %>% 
  mutate(banking_crisis = as.integer(if_else(banking_crisis == "crisis", 1, 0))) %>% 
  glm(banking_crisis ~ ., data = .) %>% 
  broom::tidy()

dat %>% 
  mutate(banking_crisis = as.integer(if_else(banking_crisis == "crisis", 1, 0))) %>% 
  glm(banking_crisis ~ year + exch_usd + inflation_annual_cpi + independence, data = .) %>% 
  broom::tidy()


```

