---
title: "heart-bagging"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
require(tidyverse)
require(tidymodels)
require(broom)

heart <- read_csv("data/heart.csv") %>% 
  mutate(target = as.factor(target)) %>% 
  mutate(id = row_number())

skimr::skim(heart)
```

1) age - age in years
2) sex - 0 = f; 1 = m
3) cp - chest pain type (4 values)
4) trestbps - resting blood pressure
5) chol - serum cholestorol in mg/dl
6) fbs - fasting blood sugar > 120 mg/dl
7) restecg - resting electrocardiographic results (values 0,1,2)
8) thalach - maximum heart rate achieved
9) exang - exercise induced angina
10) oldpeak - ST depression induced by exercise relative to rest
11) slope - the slope of the peak exercise ST segment
12) ca - number of major vessels (0-3) colored by flouroscopy
13) thal - 0 = normal; 1 = fixed defect; 2 = reversible defect
14) target - 0 = less chance of heart attack 1 = more chance of heart attack

## some simple models
## skipping normality checks

```{r}
# logistic regression
heart %>% 
  glm(target ~ age + sex, 
      data = ., 
      family = "binomial") %>% 
  tidy()
# for every one unit increase in Xi, there is a ___ increase in the log odds of being at elevated risk for heart attack
# for reference, a log odds of -1.39 corresponds to a probability of 0.2
```

## some more robust modeling
## 1. recipe

```{r}
set.seed(8765)

# create training/test 75/25 split
heart_split <- initial_split(heart, strata = "target")
heart_train <- training(heart_split)
heart_test <- testing(heart_split)

# data pre-processing 
heart_rec <- recipe(target ~., data = heart) %>% 
  update_role(id, new_role = "id") %>% 
  step_normalize(all_predictors())

# finalize recipe
heart_prep <- prep(heart_rec)
```

## peak at prepped data

```{r}
juice(heart_prep) %>% 
  head()
```

## 2. model

```{r}
# package for bootstrap aggregating
# type of model averaging
# "Bagging is used with decision trees, where it significantly raises the stability of models in the reduction of variance and improving accuracy, which eliminates the challenge of overfitting."
# - corporatefinanceinstitute.com
require(baguette)

# create workflow object
wf <- workflow() %>% 
  add_recipe(heart_rec)

# Tree-based model  
tree_spec <- bag_tree() %>% 
  set_engine("rpart", times = 100) %>% 
  set_mode("classification")

# Multivariate Adaptive Regression Splines
mars_spec <- bag_mars() %>% 
  set_engine("earth", times = 100) %>% 
  set_mode("regression")

# doParallel::registerDoParallel()

tree_res <- wf %>% 
  add_model(tree_spec) %>% 
  fit(heart_train)

require(vip)

# tree_res %>% 
#   pull_workflow_fit() %>% 
#   vi(method = "model")

mars_res <- wf %>% 
  add_model(mars_spec) %>% 
  fit(heart_train)

```

## evaluate

```{r}
test_res <- heart_test %>% 
  bind_cols(predict(tree_res, heart_test)) %>% 
  rename(.pred_tree = .pred_class) %>% View()
  bind_cols(predict(mars_res, heart_test)) %>% 
  rename(.pred_mars = .pred_class)

test_res %>% 
  metrics(target, .pred_tree)

test_res %>% 
  metrics(target, .pred_mars)
```

## make new observations and predict target

```{r}
new_pats <- crossing(
  age = sample(heart_test$age + 5, 2, replace = TRUE),
  sex = sample(c(0, 1), 2, replace = TRUE),
  cp = sample(c(1:3), 2, replace = TRUE),
  trestbps = seq(100, 150, by = 10),
  chol = seq(150, 500, by = 50),
  fbs = sample(c(0, 1), 2, replace = TRUE),
  restecg = sample(c(0, 1), 2, replace = TRUE),
  thalach = seq(125, 180, by = 15),
  exang = sample(c(0, 1), 2, replace = TRUE),
  oldpeak = seq(0, 3, by = .5),
  slope = sample(c(0:3), 2, replace = TRUE),
  ca = sample(c(0:3), 2, replace = TRUE),
  thal = sample(c(0:3), 2, replace = TRUE)
) %>% 
  mutate(id = row_number())

new_preds <- new_pats %>% 
  bind_cols(predict(tree_res, new_pats))

new_preds %>% 
  ggplot(aes(age, chol, color = .pred_class)) +
  geom_line(size = 1.2, alpha = 0.7) +
  geom_point(size = 1.5)

```

## PCA

```{r}
heart_rec <- recipe(target ~., data = heart) %>% 
  update_role(id, new_role = "id") %>% 
  step_normalize(all_predictors()) %>% 
  step_pca(all_predictors())

pca_prep <- prep(heart_rec)

# how many PCs to keep?
# maybe just 2?
summ <- helpRs::get_pca_summary(pca_prep)

summ

require(broom)

tidy_pca <- tidy(pca_prep, 2)

tidy_pca %>% 
  filter(component %in% paste0("PC", 1:5)) %>% 
  mutate(component = fct_inorder(component)) %>% 
  ggplot(aes(value, terms, fill = terms)) +
  geom_col(show.legend = FALSE) +
  facet_wrap(~component, nrow = 1) +
  labs(y = NULL)

require(tidytext)

tidy_pca %>% 
  filter(component %in% paste0("PC", 1:4)) %>% 
  group_by(component) %>% 
  top_n(8, abs(value)) %>% 
  ungroup() %>% 
  mutate(terms = tidytext::reorder_within(terms, abs(value), component)) %>% 
  ggplot(aes(abs(value), terms, fill = value > 0)) +
  geom_col() +
  tidytext::scale_y_reordered() +
  facet_wrap(~component, nrow = 1, scales = "free_y") +
  labs(y = NULL, fill = "Positive?")

juice(pca_prep) %>% 
  ggplot(aes(PC1, PC2, label = id)) +
  geom_point(aes(color = target), size = 2) +
  # geom_text(check_overlap = TRUE, hjust = "inward") +
  labs(color = NULL)

juice(pca_prep) %>% 
  ggplot(aes(PC1, PC3, label = id)) +
  geom_point(aes(color = target), size = 2) +
  # geom_text(check_overlap = TRUE, hjust = "inward") +
  labs(color = NULL)

juice(pca_prep) %>% 
  ggplot(aes(PC2, PC3, label = id)) +
  geom_point(aes(color = target), size = 2) +
  # geom_text(check_overlap = TRUE, hjust = "inward") +
  labs(color = NULL)
```

## Use principal components as independent variables in new linear model

```{r}
juice(pca_prep) %>% 
  glm(target ~PC1 + PC2 + PC3 + PC4 + PC5, data = ., family = "binomial") %>% 
  tidy() %>% 
  filter(p.value <= 0.05) %>% 
  arrange(desc(estimate))
```






